Meta-analytic approaches and effect sizes to account for 'nuisance heterogenity' in comparative physiology

### 2021-07-22 MS/ms.Rmd
updated to ms and cache. Mainly on Table 1

There are many phenomena driving effect heterogeneity. First, there is inherent uncertainty in estimating the ‘true’ population effect (e.g., a correlation coefficeiont) from smaller samples [@Borenstein2019]. A formal meta-analysis weights sample estimates by their precision, and removes this source of heterogeneity [@Gurevitch2018; @Koricheva2013]. Removing sampling variance is possible because we know, mathematically, how to calculate it for many effect measures. While unweighted analyses are common, weighting studies makes overall estimates more precise, and less susceptible to publication bias (see below), even if the overall mean itself is unbiased [@Morrissey2016a]. Second, there are sociological factors impacting heterogeneity. These include the ease with which novel and significant results are published relative to non-significant results, manifesting as publication bias [@Jennions2013; @Nakagawa2021b; @Rothstein2005]. Third, real biological processes can drive effect heterogeneity. This is particularly true in comparative physiology where we synthesise data from different species with varying life-histories, habitats, mating systems and reproductive modes. As comparative physiologists, this is the stuff that gets us up every morning! Some variation could result from shared evolutionary history (i.e., phylogeny) or because of the different thermal environments species (or populations) occupy. We can model these biological factors to test predictions from hypotheses that we believe are key players impacting mean effect size and direction. This improves our understanding of the biological world around us and helps inform on future research directions.

@@

We tend to agree with @Glass2015, and his sentiment is particularly relevant to comparative physiology. For example, if we want to understand the impact of a pollutant on reproduction in aquatic organisms, including aquatic mammals and fish would provide answers to this general question, even if their reproductive biology and physiology are different. Synthesizing diverse study outcomes allows us to also understand critical features of the literature, study systems (e.g., different effects of pollutants on fish and mammal reproduction), and approaches (e.g., short-term vs long-term studies) that can explain the diversity of effect outcomes observed. Doing so provides a rich set of conclusions that can draw attention to important sources of variability that might have otherwise been overlooked. Having said that, it is still very important to understand the limitations and diversity of studies included in a meta-analysis, and how this can affect interpretations made -- an important reason why transparency and reproducibility are such a prominent feature of meta-analysis (or ideally should be) [@ODea2021]. The validity of general conclusions to certain questions (e.g., what is the overall effect of a pollutant on reproduction?) will inevitably depend on the homogeneity of effects being synthesized. If the estimate of an overall effect coincides with low heterogeneity then the effects are reasonably consistent [@Borenstein2019]. If, however, an overall effect is accompanied by large heterogeneity, then we would shift focus and explore drivers of that heterogeneity. For example, given the very different physiology of aquatic mammals and fish, understanding an overall combined effect on reproduction may not be particularly relevant and may even be misleading. It may, therefore be more useful tounderstand overall effects in each of these groups separately, and assess the extent to which they differ.

@@

# I'm just listing a bunch here, but we should think about what we want to include. Here's an example of how we can make a table with all equations. Fun! Really useful website here: https://math.meta.stackexchange.com/questions/5020/mathjax-basic-tutorial-and-quick-reference. Something wrong with character coding when rendering word document. Has to do with the log odds ratio equation. Renders fine in html, but just stops when word
names <- c("Mean",
           "Log Standard deviation, \\textit{lnSD}",
           "Log Response Ratio, \\textit{lnRR}",
           # leave this for now??? - we introduce this later
           "Standardised Mean Difference, \\textit{SMD}",
           "\\textit{Zr} (Fisher Transformation of \n Correlation Coefficient, r)")
           #"Log Odds Ratio",
           #"")

@@

# Note: It is esssential that you be clear that the function 'compose' is exported from the flextable package. That's because purr also has the same function, so it gets messy.
tablea <- flextable(table_fin)                                                                                %>% 
          flextable::compose(j = 1:4, part = "header", value = as_paragraph(as_b(c("Effect Measure", "Definition", "Sampling Variance", "Examples")))) %>% 
          flextable::compose(j = 2:3, value = as_paragraph(as_equation(c(eqns_yi, eqns_vi))))             %>%
          flextable::compose(j = 4, value = as_paragraph(eg))             %>%
          flextable::width(j = 1, 3)                                                                          %>% 
          flextable::width(j = 2, 2) %>% 
          flextable::width(j = 4, 5) %>% 
          flextable::footnote(i = 4, j = 1, value = as_paragraph(as_equation(footnotes)),
               ref_symbols = c("a"),
               part = "body", inline = TRUE)

>>>>

There are many phenomena driving effect heterogeneity. First, there is inherent uncertainty in estimating the ‘true’ population effect (e.g., a correlation coefficeiont) from smaller samples [@Borenstein2019]. A formal meta-analysis weights sample estimates by their precision, and removes this source of heterogeneity [@Gurevitch2018; @Koricheva2013]. Removing sampling variance is possible because we know, mathematically, how to calculate it for many effect measures. While unweighted analyses are common, weighting studies makes overall estimates more precise, and less susceptible to publication bias (see below), even if the overall mean itself is unbiased [@Morrissey2016a]. Second, there are sociological factors impacting heterogeneity. These include the ease with which novel and significant results are published relative to non-significant results, manifesting as publication bias [@Jennions2013; @Nakagawa2021b; @Rothstein2005]. Third, real biological processes can drive effect heterogeneity. This is particularly true in comparative physiology where we synthesise data from different species with varying life-histories, habitats, mating systems and reproductive modes. As comparative physiologists, this is the stuff that gets us up every morning! We can model these biological factors to test predictions from hypotheses that we believe are key players impacting mean effect size and direction. This improves our understanding of the biological world around us and helps inform on future research directions.

@@

We tend to agree with @Glass2015, and his sentiment is particularly relevant to comparative physiology. For example, if we want to understand the impact of a pollutant on reproduction in aquatic organisms, including aquatic mammals and fish would provide answers to this general question, even if their reproductive biology and physiology are different. Synthesizing diverse study outcomes allows us to also understand critical features of the literature, study systems (e.g., different effects of pollutants on fish and mammal reproduction), and approaches (e.g., short-term vs long-term studies) that can explain the diversity of effect outcomes observed. Doing so provides a rich set of conclusions that can draw attention to important sources of variability that might have otherwise been overlooked. Having said that, it is still very important to understand the limitations and diversity of studies included in a meta-analysis, and how this can affect interpretations made -- an important reason why transparency and reproducibility are such a prominent feature of meta-analysis (or ideally should be) [@ODea2021]. The validity of general conclusions to certain questions (e.g., what is the overall effect of a pollutant on reproduction?) will inevitably depend on the homogeneity of effects being synthesized. If the estimate of an overall effect coincides with low heterogeneity then the effects are reasonably consistent [@Borenstein2019]. If, however, an overall effect is accompanied by large heterogeneity, then we would shift focus and explore drivers of that heterogeneity. For example, given the very different physiology of aquatic mammals and fish, understanding an overall combined effect on reproduction may not be particularly relevant and may even be misleading. It may, therefore be more useful to understand overall effects in each of these groups separately, and assess the extent to which they differ.

@@

# I'm just listing a bunch here, but we should think about what we want to include. Here's an example of how we can make a table with all equations. Fun! Really useful website here: https://math.meta.stackexchange.com/questions/5020/mathjax-basic-tutorial-and-quick-reference. Something wrong with character coding when rendering word document. Has to do with the log odds ratio equation. Renders fine in html, but just stops when word
names <- c("Mean",
           "Log Standard deviation, lnSD",
           "Log Response Ratio, lnRR",
           # leave this for now??? - we introduce this later
           "Standardised Mean Difference, SMD",
           "Zr (Fisher Transformation of \n Correlation Coefficient, r)")
           #"Log Odds Ratio",
           #"")

@@

# Note: It is esssential that you be clear that the function 'compose' is exported from the flextable package. That's because purr also has the same function, so it gets messy.
tablea <- flextable(table_fin)                                                                                %>% 
          flextable::compose(j = 1:4, part = "header", value = as_paragraph(as_b(c("Effect Measure", "Definition", "Sampling Variance", "Examples")))) %>% 
          flextable::compose(j = 2:3, value = as_paragraph(as_equation(c( eqns_yi, eqns_vi))))             %>%
          flextable::compose(j = 4, value = as_paragraph(eg))             %>%
          flextable::width(j = 1, 3)                                                                          %>% 
          flextable::width(j = 2, 2) %>% 
          flextable::width(j = 4, 5) %>% 
          flextable::footnote(i = 4, j = 1, value = as_paragraph(as_equation(footnotes)),
               ref_symbols = c("a"),
               part = "body", inline = TRUE)
